Parsing mPFC: Community analysis

### 2018-08-28 Manuscript.Rmd
Change of plan. Going back to a regular markdown doc, not rticle

## Introduction
One of the main goals of human cognitive neuroscience is to be able to prescribe function to brain regions, either in isolation or in their communication with others. Among general regions of the brain, the medial prefrontal cortex (mPFC, located in the midline frontal portion of the brain) has been suggested to subserve the cognitive abilities that differentiate humans from other animals (Barbas & Garcia-Cabezas, 2016). Specifically, this region has been related to a variety of functions, including decision making (Bartra, McGuire, & Kable, 2013), memory encoding (Shacter, Addis, & Buckner, 2007), and default mode deactivation (DMN) (Yeo et al., 2011). While much has been discovered about this area in primates (Barbas & Garcia-Cabezas, 2016), the lack of direct measurements of neuronal activity makes this a tough area to record in humans, one that nowadays relies mostly on functional MRI (fMRI). Unfortunately, the mPFC is especially challenging to image, as the oxygen-related signal recorded by fMRI is contaminated by the oxygen in the sinuses (Logothetis, 2008). In addition, this area is subject to more idiosyncratic cortical folding than any other region (Zilles, Palomero- Gallagher, & Amunts, 2013), thus adding a level of complexity to the generalizable dissection of topographic functional roles within mPFC.

Studies of valuation have consistently shown functional overlap with DMN. Recent meta-analytic work has attempted to distinguish the brain areas belonging to each by examining the locus of peak activity across studies related to the behavioral features mentioned above (Bartra et al., 2013; Acikalin, Gorgolewski, & Poldrack, 2017). While this approach offers valuable information, it is subject to the quality of imaging in each study, as well as their corresponding differences in tasks used to examine choice. In addition, recent findings suggest that the functional topology of the brain is better understood by accounting for organizational idosyncrasies (Braga and Buckner, 2018; Yeo et al., 2018, Gordon et al., 2017, Gratton...).

Given the shortcomings of study-specific and cross-study analyses mentioned above, there is potential value in combining complementary strengths from both approaches. While my lab is primarily interested in the portion of mPFC dedicated to decision making, in order to bypass the heterogeneity in behavioral tasks, it might be best to disentangle the mPFC by focusing first on which of its sub-regions belong to resting-state DMN, and which do not. The advantage lies in that (1) all resting state scans are performed equally (the subject lies in the scanner with their eyes open), and (2) there is a high quality resource of resting state data provided by the Human Connectome Project. Focusing on resting state data has been successful in determining a number of neural phenomena, including disentangling auditory and visual attention areas (Tobyne, Osher, Michalka, & Somers, 2017).

This approach opens an opportunity to examine brain dynamics. For example, a region’s function might depend on its connectivity at any given time with one brain community or another. An ever-growing trend in human neuroimaging has been to apply network statistics to tackle these functional challenges. One great advantage of network science is the ability to characterize connectivity in a way that allows for comparisons among individuals in a meaningful way (Garcia, Ashourvan, Muldoon, Vettel, & Bassett, 2018), thus letting us address the folding idiosyncrasies mentioned above. Therefore, I propose to use some of the statistics that we have learned in class to estimate intrinsic functional subdivisions of the mPFC at the individual subject level. Particularly, I would like to distinguish communities related to DMN from non-DMN, so that we can generate more informed topographic targets for future studies of decision making.

In this study, I try to address my question at 3 levels of granularity: 1) across the literature (meta-analysis), selecting a number of regions of interest (ROI) that show significant overlap and specificity between DMN and decision making; 2) full-brain connectivity of actual fMRI data, using a pre-specified brain atlas; and 3) characterizing the connectivity of surface vertices (small units of function that contain oxygen-dependent neural information across time) for the regions selected in step one. I performed the last two steps at an individual subject level for 3 subjects, hoping to find some commonalities among them. In this way, I attempted to extract functional topographical information from topological network features.

### scrabbling
Previous meta analytic studies have provided important insights on the psychological phenomena attributed to mPFC subregions. For example, ADV et al find subdivisions of mPFC and MCC that are significantly tied to decision making in the literature. In particular, they show that more ventral regions are more tightly associated with reward and fear. However, this pattern overlaps with their finding that vmPFC also coactivates with DMN regions, and find no topographical distinction between these phenomena. Similarly, the results from Bartra et al and Laird et al show that this region can't be subdivided into DMN vs subjective value (acikalin, chris, russ). However, meta analysis rely on peaks often derived from group-averaged data. Recent work has prescribed relevance to the analysis of single subjects in fMRI. Yeo et al (2018) propose the use of individualized topologies when identifying functional networks. Gordon et al identify important idiosincrasies in oversampled subjects. Similarly, Braga and Buckner show that finer subdivisions of the DMN are possible when looking at each subject independently. This opens up the possibility that distinctions between DMN and subjective value are possible at this finer level of resolution, but would otherwise be masked in meta analytic work.
Braga & Buckner's work is particularly informative when it comes to subdividing DMN. However, the seed-based functional connectivity analysis is sensitive to the selection of the seed. A natural next step would be to recruit all possible patterns of covariation within this network. A popular alternative is to rely on graph theoretic methods, which provide a characterization of networks that organically fits brain dynamics (provides an organic avenue to understand brain dynamics?), and thus permits the consideration of the full covariance structure of brain networks. (ADV & Tal somewhere here).

add tors work too

It is true that subregions of mPFC can be multifaceted, and ADV et al do suggest that given that no single psych construct belonged to each area. That said, the possibility remains that this is a feature of group averaging.

## Methods
**Meta-analysis**
To perform the first step, I used data from a metanalysis that gathered peak brain coordinates of activity from studies of valuation (n = 27 studies) and DMN (n = 77). These are the surviving areas post-statistical thresholding from each study. I analyzed this data by mapping the peak cortical coordinates to an atlas of the human cortical surface (Glasser et al., 2016). This produced a list of standardized parcels that were reported on each study. The advantage of utilizing this atlas was threefold: first, it reduced the number of functional vertices from 60,000 (cortical surface vertices) to 360 (cortical parcels). Second, it standardizes the vertices from the first to the second stage of my study. Finally, it allowed me to project each area on a brain space without much clutter. Since each vertex is embedded with topographical information, and since functional topography is essential to my question, I mostly project my results onto brain space.

I thus generated two undirected, loop-less, weighted graphs in which each vertex was a brain parcel, and edges correspond to the number of times each pair of them co-occurred across studies. Since different decision tasks use different sensory modalities, and since sensory areas get sometimes reported as unrelated peak coordinates in DMN studies, I bootstrapped the co- occurrences across studies and removed edges whose co-occurrences lied below the 90th percentile. Based on this, only edges with co-occurrences higher than 8 were preserved, and the remaining unconnected vertices (brain regions) were discarded. With these graphs set up, I computed the strength and betweenness centrality in order to capture each parcel’s centrality in the valuation and DMN literatures separately.

In order to select of regions of interest (ROI) from mPFC and other brain areas that are central to valuation and DMN literatures, I first bootstrapped the mean strength for each graph separately, and chose areas with strength above the 95th percentile of this distribution. Next, I took the intersection of these areas, such that the final ROIs were brain regions that overlapped between DMN and valuation in the literature (and thus my main target for dissociation).

**Whole Brain Parcellated connectivity**
For the second step, in order to actually quantify the intrinsic connectivity of these ROIs in the brain, I used data from the Human Connectome Project. Specifically, I acquired resting-state fMRI activity from 3 subjects, all of whom have been preprocessed according to industry standards to allow for group-level comparisons (Tobyne et al., 2017). Each data set contained around 1 hour of signals represented as 4800 time points, and each subject contained 60,000 vertices. I parcellated each subject’s brain according to the atlas mentioned above, so that each parcel contained the mean time series from its vertices, and correlated the mean time series among all parcels (Pearson). This produced a weighted adjacency matrix for each subject, where each of the 360 parcels is a vertex, and edges the correlations among all of them. Next, I took the exponential of the correlations, Fisher transformed them (i.e. tanh) so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant (FDR < 0.05). Again, I quantified each parcel’s strength and betweenness, averaged them across subjects, and examined whether any of the ROIs showed uniquely high centralities. For this, I performed a permutation comparing the mean strength between the ROIs and the rest of the brain for each participant.

@@

## Introduction

This report focuses on the community detection results, but it might be worth adding the meta-analytic steps here eventually as well. In general, the similarity matrices show a representative subsample of subjects (n = 10) to conserve space (unless noted), but the boxplots and formal tests include all subjects. The trends shown by the example subjects generalize to the larger sample anyways.

## Partition Comparison: Full Datasets

@@

Finally, we examined whether these topological features were defined by the structure of the brain by correlating each subject's Fiedler vector with their structural MRI data. We found that the Fiedler vector is uncorrelated with myelin density (mean correlation = `r round(meanMyelin, digits = 2)`; SD = `r round(sdMyelin, digits = 2)`), cortical folding (mean correlation = `r round(meanCurv, digits = 2)`; SD = `r round(sdCurv, digits = 2)`), and surface thickness (mean correlation = `r round(meanThick, digits = 2)`; SD = `r round(sdThick, digits = 2)`). While these results suggest that each subject's functional topology is mostly independent of brain structure, we will perform more detailed analyses of this in the future (i.e. are there any outstanding structural features in a subject that match their unique topology?).

>>>>

## Introduction
  Studies of decision making have consistently associated activity in ventromedial prefrontal cortex (vmPFC) with the subjective value assigned to decision outcomes (Bartra, McGuire, & Kable, 2010; Clithero & Rangel, 20??; Hiser & Koenigs, 2017; Kable & Glimcher, 2007; Levy et al., 2011). However, the distributed cortical valuation system also includes posterior cingulate and more extended mPFC regions, which collectively show significant spatial overlap with the default mode network (DMN) along the medial wall (Fox et al., 2005; Laird et al., 2009; Yeo et al., 2011). This shared neural coverage has prompted the question of whether (and to what extent) these two systems can be dissociated.

  Previous meta-analytic work has provided important insights on the psychological phenomena attributed to mPFC subregions. Independent coordinate-based meta-analyses have found mirroring spatial coverage in mPFC for valuation (Bartra et al., 2010) and DMN (Laird et al., 2009). Whole brain analyses have shown that while more ventral mPFC regions are increasingly associated with reward-related elements of decision making, vmPFC strongly coactivates with DMN regions across the literature, with no clear topographical distinction between these phenomena (de la Vega et al., 2016). More focused meta-analytic work aimed at disentangling DMN from subjective value has determined that these systems are inseparable within mPFC (Acikalin, Gorgolewski, & Poldrack, 2017). These findings highlight the multifaceted embedding of psychological constructs in subregions of mPFC, and suggest that DMN and valuation might indeed be subserved by the same system. However, recent activation-based meta-analytic work has demonstrated considerable domain specificity in this area (Kragel et al., 2018). While informative, a key limitation in distinguishing constructs through this approach is that meta-analyses often rely on information derived from group-averaged data. Averaging has traditionally been beneficial in identifying tendencies in brain function and organization when faced with short scanning sessions per subject. However, it has been shown that mPFC is subject to considerable idiosyncratic cortical folding (Zilles, Palomero- Gallagher, & Amunts, 2013) and inter-subject functional variability (Mueller et al., 2013) compared to the rest of the brain, thus making group-averaging studies difficult to interpret.

  Seeking to remedy these shortcomings, recent work has prescribed relevance to the analysis of single subjects in fMRI. Individual analyses of fMRI data have identified idiosyncratic, reliable, and valid functional organization that would otherwise be blurred in aggregative estimations (Laumann et al., 2015; Gordon et al., 2017; Gratton et al., 2018). Furthermore, subject-specific network arrangements have been found to predict behavioral characteristics (Yeo et al., 2018). In regards to DMN, the trend of individualized analyses has led to finer idiosyncratic subdivisions of the DMN through careful selection of seed-based functional connectivity (Braga & Buckner, 2018). It is thus possible that the indisitinguishable overlap of DMN and valuation effects can be attributed to a lack of spatial resolution that is better understood at the individual level. With this in mind, an important first step in disentangling these phenomena would be to determine the degree of topological heterogeneity of the DMN present within mPFC.

  A persistent issue in analyzing mPFC in individuals is the various factors inducing signal dropout in its more ventral segments (Logothetis, 2008). As such, there is value in examining all possible sources of covariation that compose DMN, instead of estimating topologies one seed at a time. Connectome-based analyses of resting state functional connectivity (rsFC) have been fruitful in characterizing individualized functional topologies that match task-induced activity (Gordon et al., 2017; Laumann et al., 2015; Tobyne et al., 2018). A popular approach to analyze these connectomes is to rely on graph theoretic methods, which provide an organic avenue to understand brain network dynamics (Bassett, Zurn, & Gold, 2018). From these, community detection algorithms have been successfully utilized to section brain networks into cohesive substructures (Garcia, Ashourvan, Muldoon, Vettel, & Bassett, 2018). Such communities represent clusters of network nodes that are more connected with each other than with the rest of the network (Fortunato & Hric, 2016). Among the algorithms used to estimate these communities, modularity has been widely effective in subsectioning brain networks into multiple groups (Garcia et al., 2018). However, in attempting to refine the DMN topology at the individual level, dividing these regions into what belongs and does not belong to DMN makes dissection through spectral partitioning a more valuable option. In short, spectral partitioning performs eigenvalue decompositon on an adjusted form of a correlation matrix (Laplacian matrix). The resulting eigenvector corresponding to the second-to-last eigenvalue (i.e. the Fiedler Vector) provides values to divide the network into a positive and a negative community (Fiedler, 1975). This method can thus capture a refined group of brain regions that more cohesively share activation patterns during rest.

  In this study, we aim to subsection regions commonly attributed to both DMN and subjective value into subject-specific DMN and non-DMN partitions, quantifying the degree of topographical heterogeneity present in this network within and across individuals. This description will allow researchers to generate more precise topographic targets for future studies of decision making. We do this by capitalizing on the respective strengths of meta-analytic and subject-specific analyses of brain data. First, we define regions of interest (ROI) by identifying overlapping activation throughout the literature projected to an established brain atlas (Glasser et al., 2013). We then produce a rsFC networks of all the surface vertices within these ROIs for each individual resting state dataset from the Human Connectome Project (HCP; citation), and estimate the extent of DMN coverage through spectral partitioning. The resulting patterns are then evaluated for consistency across methods, throughout each individual's time series, and across subjects. Finally, we show some interesting properties of SP related to the stability of brain networks.

## Methods
Tentative outline:

- Meta-analysis
- fMRI data
- Network definition
- Community detection
- Evaluation

**Meta-analysis**
We used data from metan-alyses that gathered peak brain coordinates of activity from studies of valuation (Bartra et al., 2010) (n = 27 studies) and DMN (Laird et al., 2009) (n = 77). These represent the surviving areas post-statistical thresholding from each study. For each peak volumetric coordinate, 10mm sphere masks were generated and then projected to a standard cortical mesh (fs_average, 32,000 vertices) using FreeSurfer. The sphere sections that overlapped with the cortex (should I add the details of the overlap?) were mapped to an atlas of the human cortical surface (Glasser et al., 2016). This produced a list of standardized parcels that were reported on each study.

With these lists, we identified the areas that were more consistently shared between DMN and valuation. For each literature, we summed the number of times each parcel was reported across studies, and performed a one-way chi-squared test of proportions on each parcel with a null probability equivalent to the bootstrapped (5000 iterations) mean of each literature's report counts. In other words, we checked which regions were significantly (p < 0.05) more reported in their respective literatures than their counterparts. Regions of interest were then determined based on which of the remaining regions were common across literatures.

Next, we identified areas that were solely related to either literature. We limited the search space to the parcels that were contained within the DMN parcellation defined by Yeo et al. (2011). We permuted the label of every study (i.e. DMN or valuation) 5000 times while preseving the total number of studies observed in each domain, and on each iteration stored the maximum statistic from a ROI-wise chi-squared test of proportions on the number of prmuted studies that had reported each ROI. This gave us a null distribution of chi-squared values that was used to determine which ROIs were more significantly prevalent on each literature.

**fMRI Data**
In order to quantify the intrinsic connectivity of these ROIs in the brain, we acquired resting-state fMRI data from the Human Connectome Project (HCP; Van Essen et al., 2013; www.humanconnectome.org) Q6 release (n = 80, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state sessions (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) were each comprised of 1200 TRs for a total scan time of 14 min 33 s, with left-right and left-right phase encoding implemented on each day. We used this scanning structure to estimate each subject's topology based on session, day, and overall data (4800 TRs).

Beyond the in-house minimally preprocessed pipeline from HCP (Glasser et al., 2013), which include considerable motion correction, temporal denoising, highpass filtering (0.0005 Hz threshold), and MNI152 normalization, all scans went through a number of additional refinements (Tobyne et al., 2017). Only subjects with both left-right and right-left phase encoding were included. In addition, only datasets with either low motion levels (under 1.5mm) or 0.5mm mean framewise displacement (FD) were used. Volumes that displayed an FD of over 0.5 mm were considered as spikes and removed from the subject's dataset, and subjects whose spike counts were above 5% of their total data were excluded from analyses. Briefly, each dataset was bandpass filtered (DEETS HERE), and had the mean greyordinate signal regressed from every vertex time series (MGSR). Each subject's brain was comprised of 32k vertices per hemisphere. We retained only the cortical surfaces, which resulted in 59,412 total surface vertices per subject. Resting state data was acquired over four scanning sessions that occurred over the course of two days (two sessions per day).

**Within-ROI Connectivity**
To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,942 per subject) and computed the Pearson correlation of the time series of every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we took the exponential of the correlations, Fisher transformed them (i.e. tanh) so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05).

@@

## Results
## Partition Comparison: Full Datasets

@@

Finally, we examined whether these topological features were defined by the structure of the brain by correlating each subject's Fiedler vector with their structural MRI data. We found that the Fiedler vector is uncorrelated with myelin density (mean correlation = `r round(meanMyelin, digits = 2)`; SD = `r round(sdMyelin, digits = 2)`), cortical folding (mean correlation = `r round(meanCurv, digits = 2)`; SD = `r round(sdCurv, digits = 2)`), and surface thickness (mean correlation = `r round(meanThick, digits = 2)`; SD = `r round(sdThick, digits = 2)`). While these results suggest that each subject's functional topology is mostly independent of brain structure, we will perform more detailed analyses of this in the future (i.e. are there any outstanding structural features in a subject that match their unique topology?).  
## Discussion
For clinical implications, see Hiser & Koenigs 2018. In short, being able to identify functional idiosyncrasies without much data can yield important insights in diagnosis (meh, we always say that..) of a variety of psychological diseases associated with mPFC. It would be more interesting to identify the repercussions to damage in mPFC per individual. They mention that subgenual vmPFC is a common target for deep brain stimulation (or TMS) as a depression treatment. The refinement of the technique is thus coupled to being able to identify personalized functional topologies. They list a number of results related to mPFC functional connectivity too. In the end, their suggestions for clinical implications of mPFC studies rely on the individual measure of functional expanse. (note that social-relevant areas in their meta analysis look similar to our patterns, and the emotion part to Barbas latest)

## References
